### 2016-10-13 summary.tex
10.13

\documentclass[10pt,journal,compsoc]{IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  \usepackage[nocompress]{cite}
\else
  % normal IEEE
  \usepackage{cite}
\fi
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.
%
% Note that some packages require special options to format as the Computer
% Society requires. In particular, Computer Society  papers do not use
% compressed citation ranges as is done in typical IEEE papers
% (e.g., [1]-[4]). Instead, they list every citation separately in order
% (e.g., [1], [2], [3], [4]). To get the latter we need to load the cite
% package with the nocompress option which is supported by cite.sty v4.0
% and later. Note also the use of a CLASSOPTION conditional provided by
% IEEEtran.cls V1.7 and later.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
    \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
    \graphicspath{{./pics/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  %  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/tex-archive/info/epslatex/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
\ifCLASSOPTIONcompsoc
  \usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
\else
  \usepackage[caption=false,font=footnotesize]{subfig}
\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a sans serif font rather
% than the serif font used in traditional IEEE formatting and thus the need
% to invoke different subfig.sty package options depending on whether
% compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/

% *** FLOAT PACKAGES ***
%
\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Do not use the stfloats baselinefloat ability as IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/dblfloatfix/

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and
% Axel Sommerfeldt. This package may be useful when used in conjunction with
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/url/
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

\usepackage{enumitem}
%\usepackage[]{algorithm2e}
\usepackage[]{algpseudocode}
\usepackage{algorithm}
\usepackage{multirow}
\usepackage{slashbox}
\usepackage{color}
\usepackage[numbers,sort&compress]{natbib}
%\usepackage{ulem}
%\usepackage[numbers,sort&compress]{natbib}
\renewcommand{\arraystretch}{1.3}

\usepackage{color}
\usepackage{bm}
%\usepackage{mathrsfs}
\usepackage[mathcal]{eucal}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{wrapfig}
% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\newcommand{\tabincell}[2]{\begin{tabular}{@{}#1@{}}#2\end{tabular}}

\floatname{algorithm}{Algorithm}

\begin{document}
\title{Efficient, Robust and Exact Boolean Operations on N Triangular Mesh Primitives}
\author{Rui~Wang,~
        Xudong~Jiang,~
        Hongbo~Fu,~
        Bin~Sheng,~
        Ping Li~and
        Enhua~Wu
\IEEEcompsocitemizethanks{

\IEEEcompsocthanksitem R. Wang and B. Sheng are with the Department of Computer Science and Engineering, Shanghai Jiao Tong University. Email:\{jhcz,shengbin\}@sjtu.edu.cn

\IEEEcompsocthanksitem X. Jiang is with Autodesk China Research \& Development Center. Email: xudong.jiang@autodesk.com

\IEEEcompsocthanksitem H. Fu is with the School of Creative Media, City University of Hong Kong. Email: hongbofu@cityu.edu.hk

\IEEEcompsocthanksitem P. Li is with the Department of Mathematics and Information Technology, The Hong Kong Institute of Education. Email: pli@ied.edu.hk

\IEEEcompsocthanksitem E. Wu is currently a research professor at State Key Lab. of Computer Science, Institute of Software, Chinese Academy of Sciences. Email: ehwu@umac.mo}
}

%\markboth{Journal of \LaTeX\ Class Files,~Vol.~13, No.~9, September~2014}%
%{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer Society Journals}

\IEEEtitleabstractindextext{
\begin{abstract}
  Constructive solid geometry (CSG) is widely used for computer aided design and manufacturing. However, boolean operations, which compute the boundary representation of CSG, have been suffered from robustness and efficiency problem for more than three decades. In this paper, we propose a fast and exact boolean operation approach which is theoretically robust for regular meshes. This is done by introducing plane-based geometry into boolean operations and strictly follow the principle of no geometry constructions during processing. Different from previous BSP-based methods, our method is with only one-time conversion to the plane representation in intersection areas, instead of a series of complex conversions between different representations. Also, by using geometry connectivity and space neighboring information, our method reuse the computation iteratively, saving much computation time. Comparison experiments with state-of-the-art techniques show that our method is unconditionally exact and robust, while is competitively efficient with non-robust methods and much faster than the robust ones.
\end{abstract}

\begin{IEEEkeywords}
boolean operations, triangle mesh, CSG evaluation, plane-based geometry
\end{IEEEkeywords}

}

\maketitle

\IEEEdisplaynontitleabstractindextext
\IEEEpeerreviewmaketitle

\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
\IEEEPARstart{C}{onstructive} solid geometry (CSG) has long been a popular modeling tool of computer-aided design and computer-aided manufacturing (CAD/CAM). It constructs complex models by combining primitives using a series of regularized boolean operations \cite{requicha1977mathematical,tilove1980closure}: union, intersection and difference. CSG can be converted into the widely-used boundary representation (B-rep) through boolean operations, which is an classical topic with history of over three decades.

In general, there are two major types of boolean algorithms according to how they deal with intersection between primitive meshes. Approximate methods redicretize the intersection areas, fit vertices approximatively and rearrange topology. On the other hand, exact method do not change vertex positions, retain as many input elements  (e.g., faces, vertices and topology) as possible. In many applications, exact methods are preferred for their accuracy. Also, using exact methods, the clear mapping between the surface of input and output meshes makes it easier for the inheritance of information like face colors and materials. However, for exact boolean operation methods, there is always a bargain between robustness and efficiency. Recent exact boolean operation methods try to find faster ways to compute boundaries of CSG. However, for robustness, the answers are often exact arithmetic \cite{barki2015exact,zhou2016mesh} or trivial workarounds, like setting a tolerance \cite{feito2013fast,segal1990using} or rotating the primitive with small angles to avoid coplanar cases \cite{douze2015quickcsg}. However, none of the solution is good enough. While methods using exact arithmetic are significantly slower, workarounds are often case dependent and not reliable for complex CSGs.

We develop an unconditionally robust and exact boolean operation method for triangle meshes without using exact arithmetic. The robustness of our method is based on the observations by Sugihara and Iri \cite{sugihara1990solid} that boolean operations can be performed without geometry constructions if representations are based on planes. This idea of using plane-based representation (P-rep) has been developed by Berstein and Fussell \cite{bernstein2009fast} by combining it with binary search partition (BSP) techniques \cite{naylor1990merging,thibault1987set}. Later, Campen et. al. \cite{campen2010exact} improved the method using octrees to resolve intersections locally, avoiding inefficient large BSP tree construction. Despite of that, Campen et. al.'s method still suffers from performance problem because of the complex conversions among B-reps, P-reps and BSP representations. Also, the incoherence between BSP representation, where boolean operations actually happens, and B-reps, which are input and output meshes, requires extra steps to connect the reconstructed intersection areas with the unchanged areas, leading to extra computation burden and complex topology.

We develop a boolean operation method based on B-reps of meshes with geometry connectivity (e.g., halfedge \cite{mcguire2000half}). Additionally, we embed P-rep of faces into the B-rep, forming a hybrid representation, which allows us to take the advantage of both representations for robustness and efficiency. This hybrid representation requires only one-time conversion from B-rep to P-rep in intersection areas of meshes.
%that allows us to take advantage of geometry connectivity for classification acceleration \cite{feito2013fast,ogayar2015deferred,updegrove2016boolean}, which is hard for BSP-based methods.
Moreover, our method can perform boolean operations with more than two primitives in one call, instead of recursively doing incremental boolean operations. Despite of extra complexity, non-incremental boolean operation may save much computation time by avoiding repetitive computation. The major difficulty of embedding P-rep into B-rep is the incoherence between the two representations. Many geometry operations (e.g., coordinates projection and compute the barycenter coordinate of triangle) available on B-rep can be very complex and thus inefficient on P-rep. On the contrary, all geometry predicates have to be performed under P-rep for exactness and robustness. We solve this problem by various methods to constrain the computation within plane-based geometry predicates. The experiments show that our method is much faster than existing robust exact methods, while only about two to three times slower than the state-of-art non-robust exact methods.

\section{Related Work}

Boolean operations have been researched for over three decades since it inception in the 1980s \cite{requicha1985boolean, laidlaw1986constructive}. Previous works of boolean operations can be classified into two categories: exact methods and approximate methods. Exact methods retain vertex positions. Also the topology of the input meshes is preserved as much as possible. The coordinates of intersection points between input meshes are computed by input configurations to construct the topology around intersection. However, the coordinates cannot be represented exactly using fix-precision float-point numbers and the robustness is always a problem. On the other hand, approximate methods rediscretize the input mesh surfaces using techniques such as voxels, octrees and Layered Depth Images. They can reach better performance and robustness, but loss of geometry information and precision is inevitable.

\subsection{Exact Methods}

\iffalse
\begin{figure}[t]
\centering
\includegraphics[width=2.5in]{coplanarexample}
\caption{{\color{red}{Sketch: the example of coplanar cases in a boolean operations with four cylinders}}}
\label{fig:coplanarexample}
\end{figure}
\fi

For exact methods, some \cite{ogayar2015deferred,douze2015quickcsg,zhou2016mesh,xu2013fast,feito2013fast} pursues efficiency but cannot guarantee robustness in degenerate cases. Most of them are implemented by fix-precision float-point arithmetic and numerical errors are inevitable. Douze et al. \cite{douze2015quickcsg} implemented very efficient method that can handle very large meshes. Also, they can perform non-incremental boolean operations: evaluating the final result mesh of CSG all at once instead of decomposing the CSG into a series of incremental boolean operations. However, their method can neither deal with coplanar cases nor guarantee accuracy of topology. In fact, large CSGs are more complex and more likely to contain degenerate cases, and more vulnerable under numerical errors. Efficiency only is far from being practical.
%Also, coplanar cases can regularly happens for man-made artifact modeling (Fig. \ref{fig:coplanarexample}) and should not be treated as exceptions for boolean operation methods.

Many researchers have attempted to solve robustness issue for exact boolean operations using arbitrary precision arithmetic \cite{banerjee1996topologically, fortune1995polyhedral, keyser2004esolid, granados2003boolean, hachenberger2005boolean} and exact interval computation \cite{fang1993robustness, hu1996robust, segal1990using}. However, these methods are too expensive in computation time and memory to be practical for CSGs with massive faces. For example, the current standard of robust exact boolean operations method is CGAL's \cite{cgal:hk-bonp3-15a} exact-arithmetic implementation \cite{granados2003boolean} of Nef polyhedra \cite{bieri1988elementary}. It requires conversions between Nef polyhedra and input mesh data structures during evaluation and is more than 50 times slower than non-robust booleans \cite{bernstein2009fast}.

\subsection{Approximative Methods}

Since efficiency, accuracy and robustness are hard to be satisfied at the same time, starting from \cite{museth2002level}, some sacrifice accuracy for better efficiency and robustness. Most of such methods are based on conversion to volumetric representations. However, the quality of result mesh depends on the resolution of the volume grid and better quality requires dramatically higher resolution. To accelerate this process, some try to reduce the complexity of the output mesh \cite{varadhan2004topology} and others try to preserve non-intersected areas of the input mesh to avoid redundant tessellation \cite{pavic2010hybrid,wang2011approximate,zhao2011parallel,hable2005blister,ogayar2006gpu}. Also, with the development of general-purpose computing on graphics processing units (GPGPU), many of them utilize the grand computation power of graphics hardware for boolean operations. These methods often have good performance and are suitable for interactive applications. However, the fundamental problem of approximate methods is owing to grid-depend nature of volumetric calculations: they inevitably suffers from geometric detail loss and unwieldy topology changes.

\subsection{Plane-Based Methods}
\label{sec:pbrelated}

The concept of plane-based representation of polygonal meshes was first described by Sugihara and Iri \cite{sugihara1990solid}. P-rep provides important advantage that for boolean operations, no new primary geometry information has to be constructed to obtain the result polyhedron---it is composed of a subset of the planes of the input polyhedra. Berstein and Fussell \cite{bernstein2009fast} combined the P-rep with binary space partitioning (BSP) \cite{thibault1987set,naylor1990merging} to develop an unconditionally robust boolean operation method. Compared to volume-grid based methods, their BSP-based method increases mesh complexity more moderately and the input topology can be reconstructed better. However, it requires conversions between input data structures and P-rep which are inefficient from both storage and time views. Also, the merge of two BSP trees is an $O(mn)$ time complexity process, where $m$ and $n$ are the size of the trees. It makes this method impractical for very large input meshes. Campen and Kobbelt \cite{campen2010exact} improved their method by localization using an octree. And the refinement only takes place locally near intersections.

\section{Background and Overview}

\label{sec:overview}
Our method is designed for boolean operations on arbitrary number of regular triangle meshes \cite{requicha1985boolean}. Input meshes are required to be without self-intersecting, represented with B-reps with geometry connectivity (e.g., halfedge, winged-edge, etc.). Despite of these requirements, our method is not sensitive to topology deficiencies like holes and self-intersections and works correctly deficiencies are not near the intersection areas between meshes.

\subsection{Background}
\label{sec:paradigm}

The boolean operation of triangle meshes is in essence a process of face selection. Namely, given a boolean function, it reserves those primitive faces that pass the function and drop the others to generate the final mesh. Unfortunately, for B-rep meshes, not all the input faces can be classified as a whole, since only part of the faces belong to the final mesh. Therefore, we need an extra step aiming at detecting intersections between meshes and performing tessellation. Most of the boolean operation methods follow this two-step paradigm, which consists of intersection computation and face classification. So is our method.

During the intersection computation, input faces are tested in pairs to compute their intersections. Each input meshes are tessellated according to the intersection results, making every face be either inside, outside or on the boundary of other primitives. The tessellated meshes are what we called \emph{intersection-free meshes}. Unfortunately, under fix-precision float-point arithmetic, intersection tests are error-prone: first, degenerate cases of intersection are hard to detect; second, when there are intersections, new vertices are introduced into the geometry, whose coordinates cannot be exactly represented by fix-precision float-point numbers.

During face classification, we evaluate whether a given face $s$ belongs to the final geometry according to the n-primitive boolean function $f$ :
\begin{equation}
\lambda_f(s) = f(\boldsymbol{\Lambda}(s)) = f(\lambda_1(s), \lambda_2(s), \cdots, \lambda_n(s)).
\end{equation}
The parameter $\lambda_i(s)$ is space indicator with respect to mesh $P_i$. To compute $\lambda_f(s)$ and classify $s$, one has to know the space indicators with respect to all the primitives (denoted as $\boldsymbol{\Lambda}(s)$). A space indicator has four conditions: completely inside (\emph{in}), completely outside (\emph{out}), on the boundary with consistent normals (\emph{same}) or opposite normals (\emph{oppo}). The rules of boolean function evaluation are described in \cite{douze2015quickcsg,feito2013fast}.

\begin{figure}[t]
\centering
\includegraphics[width=2.5in]{falseclass}
\caption{In a 2D view, face $s$ is on the surface of mesh $P$. However, because we use barycenter $v_c(s)$ to compute the indicator  and the coordinates of $v_c(s)$ contain round-off error, we might falsely classify $s$ as outside of $P$.}
\label{fig:falseclass}
\end{figure}

Face classification also suffers from robustness problems. Many methods classify face according to the indicators of its barycenter, which can be computed by point-in-polyhedron test \cite{feito2013fast,campen2010exact}. However, coordinates of barycenters cannot be exactly represented and could generate false classification (Fig. \ref{fig:falseclass}). In addition, for the large amount of faces and input meshes, many methods take the benefits of the local coherence of indicators, classifying neighboring faces together \cite{pavic2010hybrid,feito2013fast,ogayar2015deferred,zhou2016mesh}. Despite of the performance improvement, it can make the robustness problem worse because the false classification may propagate to neighboring faces.

We develop an exact boolean operation method based on the two-step paradigm, being unconditional robust and as fast as possible.
\iffalse
Our method includes the following features:
\begin{itemize}
    \item Intersection computation has to avoid errors when introducing new vertices into meshes.
    \item All degenerate cases of intersection between input meshes have to be well resolved.
    \item Face classification has to be exact and robust.
    \item Acceleration using geometry connectivity is necessary for face classification.
\end{itemize}
\fi
For efficiency, the robustness of method does not rely on arbitrary precision arithmetic. Instead, we apply the plane-based techniques from BSP-based method \cite{campen2010exact}. However, BSP structure does not support geometry connectivity query in nature, which is essential for classification acceleration. Also, we do not want complex conversions between different representations. Therefore, our method is based on B-rep and we embed P-rep into intersection areas to guarantee the robust geometry computations.

\iffalse
\subsection{Intersection-Free Hybrid Mesh}

\begin{figure}[t]
\centering
\includegraphics[width=2.5in]{linkedhalfedge}
\caption{{\color{red}{Sketch: The Linked Halfedge structure}}}
\label{fig:linkedhalfedge}
\end{figure}

\begin{figure*}[!t]
\centering
\includegraphics[width=7.1in]{flowchart}
\caption{{\color{red}{Sketch: overview, will be replaced}}}
\label{fig:overview}
\end{figure*}

\label{sec:meshes}
We made an observation that the non-intersected meshes, as intermediate result of intersection computation, plays an important role of boolean operations. It is a bridge that controls the complexity of intersection computation and facilitates face classification. After the data structure of non-intersected meshes is determined, all things left are how to construct this structure and how to classify according to it.

Our non-intersected mesh, called Linked Halfedge (Fig. \ref{fig:linkedhalfedge}), is a variance of halfedge, a wide-used data structure to represent solids. Halfedge has two advantages: first, since halfedge is popular, using it as the intermediate data structure can avoid unnecessary conversion, which has the potential to have better performance and result topology; second, the geometry connectivity, by which we accelerate classification using local coherence of space indicators, can be easily retrieved through halfedge.

However, naive halfedge is far from enough. The vertex coordinates in halfedge are usually represented by float point number, which cannot exactly represent the newly introduced intersection points. Therefore, we utilize P-rep to avoid computing new coordinates. In addition, we detect all coincident vertices among input meshes (including intersection vertices). This information is used for avoiding repetitive vertices in the final mesh, reconstructing geometry connectivity, and ensuring the topology correctness.

Enlightened by Feito et. al. \cite{feito2013fast}, we found that the the topology near intersection lines between faces can benefit point-in-polyhedron test by dramatically reducing the number of faces needing traversing. Therefore, we add what we call \emph{intersection context} into the our non-intersected meshes to facilitate afterwards face classification. Intersection context identifies where an intersection lies on the surface of the related primitives. An edge of Linked Halfedge stores intersection context only if it is coincident with a certain intersection line segment between primitives. We will give the definition and usage of intersection context in details in \S\ref{sec:ir}.

To give an overall impression of Linked Halfedge, we summarize the above content and highlight the features of Linked Halfedge as follows:

\begin{itemize}
  \item Each input mesh is tessellated into a Linked Halfedge by intersection computation.
  \item Coincident vertices between different Linked Halfedge should be linked, that is, coincident vertices are shared among Linked Halfedges.
  \item If a certain edge of Linked Halfedge is coincident with intersection line segment(s), it should be associated with the corresponding intersection context(s).
\end{itemize}

\fi
\subsection{Method Overview}

There are generally four stages for our boolean method. The first three stages are to compute intersection, which can be seen as pre-processing, processing and post-processing. After that, the input meshes are tessellated into intersection-free meshes with hybrid representations. The last step is to classify each face and get the final result mesh of boolean operations.

\subsubsection{Space division}

\begin{wrapfigure}{r}[0in]{0in}
\includegraphics[width=1.2in]{octreediff}
\end{wrapfigure}

As intersection detection is performed between each pair of faces, localization is necessary to filter invalid pairs beforehand. Many space division methods can be used in this step. We use the adaptive octree for its simplicity and efficiency. Our implementation is akin to the implementation in Ogayar et al.'s method \cite{ogayar2015deferred}. Intersection between triangle faces and octree leaves are efficiently detected using the separating axis theorem \cite{gottschalk1996obbtree}. Octree leaves are classified into two types: if all faces that intersect a leaf belong to the same primitive, we call it a \emph{normal cell}. Otherwise, it is a \emph{critical cell}, within which the following intersection computation is performed.

The difference between our octree and Ogayar et al.'s is that we do not subdivide any normal cell no matter how many faces it contains. This is because subdividing normal cells benefits only the point-in-polyhedron test  \cite{frisken2002simple}, which seldom uses in our method. This simplification can save much computing time, especially when intersections between primitives are not complex and located in small regions.

\subsubsection{Triangle-triangle intersection}

Our triangle-triangle intersection test is largely based on M\"{o}ller's algorithm \cite{moller1997fast}, which is very efficient to compute intersection between two triangles. However, naive implementation of M\"{o}ller's algorithm using vertex-based geometry brings numerical errors. Therefore, we integrate P-rep into M\"{o}ller's algorithm. All intersection points and line segments are represented using plane intersections to avoid explicitly computing the coordinates. Also, we carefully deal with degenerate cases of triangle intersections, including point intersection, edge intersection and coplanar situation. Details are provided in \S\ref{section:isect}.

\subsubsection{Tessellation}

Once all intersections between triangles are detected, we need to tessellate the input meshes and construct the intersection-free meshes. In many methods like \cite{ogayar2015deferred}, constraint Delaunay triangulation (CDT) is applied to perform tessellation. However, for a CSG with more than two meshes, intersections may overlap or intersect with each other, and cannot be used for constraints of CDT directly. In addition, as our intersections are represented by planes, implementation of CDT are complex and inefficient, as most CDT algorithms (e.g. \cite{chew1989constrained}, \cite{de1992line}) is not designed to handle planes and requires explicit coordinates. Therefore, we first perform intersection refinement to resolve intersecting intersections. After that, we construct a graph-like structure suitable for P-rep intersections, called \emph{tess-graph}, to guide the exact tessellation of each face. After all faces are tessellated, we get our intersection-free hybrid meshes, where 'hybrid' means the meshes are represented by both B-rep and P-rep. Details are shown in \S\ref{sec:tessellation}.

\begin{figure}
  \centering
  % Requires \usepackage{graphicx}
  \includegraphics[width=1.7in]{sgen}\\
  \caption{Triangle-triangle intersections made by different triangle pairs may (a) overlap or (b) intersect with each other. (c) Such situation introduces extra vertex which is intersection of triangles from three different meshes.}\label{fig:p-reps}
\end{figure}

\subsubsection{Face classification}

This step is to choose faces from the intersection-free meshes to generate the final mesh. However, literally computing all the space indicators of each face is unacceptably slow for large CSG mesh. We utilize the geometry connectivity to propagate indicators in a flood-filling way, which is based on the spatial coherence of face indicators. In addition, many methods use barycenters of faces together with point-in-polyhedra test for face classification, which is not exact nor robust with fix-precision float-point arithmetics. Douze et al. \cite{douze2015quickcsg} classifies faces using space indicators of vertices, but failed to deal with coplanar conditions because of coplanar leads to ambiguity of $same$ and $oppo$ in such configurations. Different from them, we classify faces based on the classification results of exactly represented vertices, including input vertices and newly introduced intersection vertices, and carefully deal with coplanar conditions to ensure topology consistency. Details will be shown in \S\ref{sec:classification}.

\section{Plane-based Geometry}

%As we discussed in \S\ref{sec:paradigm}, one of the problem in boolean operations under vertex-based representation is it inevitably introduces new vertices, whose coordinates cannot be exactly represented using fix-precision float-point number. On the other hand, if P-rep is used, no new geometry information needs to be constructed. Therefore, we can use only geometric decision predicates to perform boolean operations.

\begin{wrapfigure}{r}[0in]{0in}
\includegraphics[width=1.2in]{p-reps}
%\caption{Plane-based representation of triangle}
\end{wrapfigure}

Using Plane-based representation, each face $s$ with $n$ edges is represented by a supporting plane $\bm{p}_{s,sp}$, where the face lies, and a list bounding planes $\{\bm{p}_{s,b}^i \ \vert\  i = 0, 1,...,n-1\}$. Each edge line $\bm{e}_s^i$ is represented by intersection $\bm{p}_{s,sp} \cap \bm{p}_{s,b}^i$. Corner vertex $\bm{v}_s^i$ is represented by intersection $\bm{p}_{s,sp} \cap \bm{p}_{s,b}^i \cap \bm{p}_{s,b}^{{(i+1)}\bmod{n}}$.

Other commonly used notations in this paper are presented here. A plane $\bm{p}$ can be represented by four scalar coefficients $a(\bm{p}), b(\bm{p}), c(\bm{p}), d(\bm{p})$. The normal of $\bm{p}$ is denoted as $\bm{n}(\bm{p})$, which is a normalized vector consisting of the first three plane coefficients. A line $\bm{l}$ can be represented by intersection of two planes $(\bm{p}_l^0 \cap \bm{p}_l^1)$, or in short, $\bm{l}\colon(\bm{p}_l^0 \cap \bm{p}_l^1)$. The positive direction of the line $\bm{l}$ is defined by $\bm{n}(\bm{p}_l^0) \times \bm{n}(\bm{p}_l^1)$. A point $\bm{v}$ can be represented by non-trivial plane triples $(\bm{p}_v^0 \cap \bm{p}_v^1 \cap \bm{p}_v^2)$, or in short, $\bm{v}\colon(\bm{p}_v^0 \cap \bm{p}_v^1 \cap \bm{p}_v^2)$.

As the precision of input coordinates is finite, we can determine in advance the upper boundary of precision needed to perform a precise decision, with which we can use fixed precision predicates to improve performance. For efficiency, we implement all geometry predicates using filtering techniques proposed by Shewchuk \cite{shewchuk1997adaptive}.

\subsection{Conversion}

\label{sec:convert}

We use Campen et al.'s method \cite{campen2010exact} for exact conversion of triangles from their vertex-based representation to plane-based representation. Given that the input vertex coordinate has $L$-bit precision including sign, the precisions of the four plane coefficients are $L_a, L_b, L_c, L_d$, which have to satisfy the following requirements:
\begin{equation}
\begin{split}
&L_a=L_b=L_c\ge 2(K-1)+1+1,\\
&L_d\ge(L_a-1)+(L-1)+2+1.
\end{split}
\end{equation}
Here $K$ is the maximum bits required to represent edge vector components in input meshes. If we use $M$ bits to store each plane coefficient, basically $M$ has to be greater than $L_d$. In our implementation, we actually use a more strong constraint $M \ge L_d+1$. Under this constraint, given a $L$-bit vertex $\bm{v_i}$, we can compute the orientation of plane $p_j$ with respect to $\bm{v_i}$ by the sign $\bm{n}(\bm{p_j})\cdot\bm{v_i} + d(\bm{p_j})$, which requires no more than $M$ bits. It allows very fast plane orientation predicates which accelerates the early rejection in our triangle-triangle intersection tests (\S \ref{section:isect}).

Let $\delta = 2^{K-L-1}$ be the relative length (in max-norm) of the longest edge in the mesh (relative to the bounding box). If we use IEEE 754 double float point number ($M=53$) and assume $\delta \le 2^{-5.5} \approx 0.022$, the maximum input precision $L$ can reach to 20, which is enough for most application (for standard IEEE 754 single float point number, the precision is 24-bit).

\subsection{Substrates}
\label{sec:substrates}
Plane-based geometry predicates used in our method include checking coincidence and co-orientation of planes, orientation of a plane with respect to a point, and whether three planes intersect in a unique point, etc. Most of them are well-discussed in previous work \cite{bernstein2009fast,banerjee1996topologically}. In the following, we focus on two advanced plane-based predicates, which are also important building blocks of our method.

\vspace{0.5em}
\noindent \textbf{Linear ordering of points}~~~~
Given an line $\bm{l}_{ab}$ with two points $\bm{v}_a\colon(\bm{p}_a^0\cap\bm{p}_a^1\cap\bm{p}_a^2)$ and $\bm{v}_b\colon(\bm{p}_b^0\cap\bm{p}_b^1\cap\bm{p}_b^2)$ on it, we need to determine the linear order of the two points along $\bm{l}_{ab}$ (Fig. \ref{fig:twopointoneline}). To solve this problem, we choose one plane not parallel with $\bm{l}_{ab}$ from the plane representation of each point. Then we convert this problem into linear order of planes and solve it using the same technique as Banerjee et al. \cite{banerjee1996topologically}. The chosen planes should have the same orientation with $\bm{l}_{ab}$ (the dot product between plane normal and $\bm{l}_{ab}$ has to be positive) and unqualified planes have to be flipped.

\vspace{0.5em}
\noindent \textbf{Circular ordering of lines}~~~~
During face tessellation, we need to know the which edges are neighboring (Fig. \ref{fig:circularorder}), that is, circular ordering of directed lines around a vertex. Lines can be sorted in a divide-and-conquer way based on the relative order of each pair of lines. Therefore, this problem is converted to that given two coplanar lines $\bm{l}_a$ and $\bm{l}_b$ with their intersection point $\bm{v}_{ab}$, compute the circular order of the two lines. Supposing the common plane of $\bm{l}_a$ and $\bm{l}_b$ is $\bm{p}_0$, we can compute the order by the sign of $\sin{\theta_{ab}}$, where $\theta_{ab}\in(-\pi,\pi)$ is the angle from $\bm{l}_a$ to $\bm{l}_b$ under the top view of $\bm{p}_0$. The sign of $\sin{\theta_{ab}}$ is the same as triple product $\bm{n}(\bm{p}_0) \cdot (\bm{l}_a\times\bm{l}_b)$,
where $\bm{n}$ means the normal direction. However, explicit computing this equation is complicated because both $\bm{l}_a$ and $\bm{l}_b$ are implicitly represented by intersection of planes.

We develop an easier way that requiring no explicit computation of $\bm{l}_a$ and $\bm{l}_b$. First, we write the P-reps of two lines as $\bm{l}_a\colon(\bm{p}_0\cap\bm{p}_a)$ and $\bm{l}_b\colon(\bm{p}_0\cap\bm{p}_b)$. Then we orthogonally decompose $\bm{n}(\bm{p}_a)$ and $\bm{n}(\bm{p}_b)$ along the normal the common plane $\bm{p}_0$:
\begin{equation}
\begin{split}
&\bm{n}(\bm{p}_a)= \bm{n}^\parallel(\bm{p}_a) + \bm{n}^\perp(\bm{p}_a)\\
&\bm{n}(\bm{p}_b)= \bm{n}^\parallel(\bm{p}_b) + \bm{n}^\perp(\bm{p}_b)
\end{split}
\end{equation}
Here superscript `$\parallel$` means the projection on $\bm{p}_s$ and `$\perp$` means orthogonal to $\bm{p}_0$. As we know $\bm{n}^\perp(\bm{p}_a)$ and $\bm{n}^\perp(\bm{p}_b)$ are both parallel to $\bm{n}(\bm{p}_0)$, we get:
\begin{equation}
\label{eq:circ1}
\bm{n}(\bm{p}_0) \cdot (\bm{n}(\bm{p}_a) \times \bm{n}(\bm{p}_b)) = \bm{n}(\bm{p}_0) \cdot (\bm{n}^\parallel(\bm{p}_a) \times \bm{n}^\parallel(\bm{p}_b)).
\end{equation}
By simple geometry reasoning, we find that the angle between $\bm{n}^\parallel(\bm{p}_a)$ and $\bm{n}^\parallel(\bm{p}_b)$ is exactly $\theta_{ab}$. So the sign of the right side of equation \ref{eq:circ1} is exactly the sign of $\sin{\theta_{ab}}$. It means we have the following nice conclusion:
\begin{equation}
\label{eq:circ2}
sign(\sin{\theta_{ab}})=  sign(\bm{n}(\bm{p}_0)\cdot(\bm{n}(\bm{p}_a) \times \bm{n}(\bm{p}_b)))
\end{equation}

Since all the three vectors on the right side are already explicitly represented, we only have to evaluate the sign of an 3x3 matrix and avoid complicated high precision float-point arithmetic.
\begin{figure}
  \centering
  % Requires \usepackage{graphicx}
  \includegraphics[width=2.5in]{twopointoneline}\\
  \caption{The configuration of linear ordering of points. Points $\bm{v}_a$ and $\bm{v}_b$ are both on line $\bm{l}_{ab}$. We convert this problem into the plane ordering between $\bm{p}_a$ and $\bm{p}_b$ along $\bm{l}_{ab}$.}\label{fig:twopointoneline}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=2.5in]{circularorder}
\caption{The configuration of circular ordering of lines. $\bm{l}_a\colon(\bm{p}_0 \cap \bm{p}_a)$ and $\bm{l}_b\colon(\bm{p}_0 \cap \bm{p}_b)$ are coplanar in $\bm{p}_0$.}
\label{fig:circularorder}
\end{figure}

\section{Exact Intersection Computation}

\label{section:isect}
In this step, intersections between faces are computed through triangle-triangle intersection test. We adopt M\"{o}ller's algorithm \cite{moller1997fast} on account of its efficiency and simplicity. However, a naive implementation of M\"{o}ller's algorithm may fail to produce correct results because of numerical errors. Therefore, we integrate plane-based geometry to make it unconditionally exact and robust. In the following, we first make a quick review of M\"{o}ller's algorithm, and then discuss the way to embed plane-based geometry. After that, we discuss how to deal with degenerate cases.

\subsection{M\"{o}ller's Vertex-Based Method}

\begin{figure}[t]
\centering
\includegraphics[width=2.5in]{projection}
\caption{$Seg_1$ is the intersection between $\bm{p}_{t_2, sp}$ and $t_1$. $Seg_2$ is the intersection between $\bm{p}_{t_1, sp}$ and $t_2$. The intersection between $t_1$ and $t_2$, which is the line segment in red, is the overlap of $Seg_1$ and $Seg_2$.}
\label{fig_projection}
\end{figure}

M\"{o}ller's algorithm computes the intersection between two triangles $t_1$ and $t_2$ in three steps as shown in Fig. \ref{fig_projection}:
\begin{itemize}[leftmargin=0.45cm]
\item[1)] An early rejection is performed by testing whether $t_1$ intersects $\bm{p}_{t_2, sp}$, the supporting plane of $t_2$. The same test is also done between $t_2$ and $\bm{p}_{t_1, sp}$.
\item[2)]The intersection between $t_1$ and $\bm{p}_{t_2, sp}$, denoted as $Seg_1$, and the intersection between $t_2$ and $\bm{p}_{t_1, sp}$, denoted as $Seg_2$, are separately computed .
 \item[3)]The intersection between $t_1$ and $t_2$ is determined by computing the overlap between $Seg_1$ and $Seg_2$ .
\end{itemize}

The non-robustness of this algorithm is from computing the coordinates of intersection vertices, which is the end points of $Seg1$ and $Seg2$. Although implementation with arbitrary precision arithmetic produces exact coordinates, it is too costly for boolean operations of large CSG. We use plane-based geometry to solve this problem by implicitly representing intersections with planes.

\subsection{Plane-Based Intersection}

\label{sec:embed}
According to the method described in \S\ref{sec:convert}, triangle faces $t$ are firstly converted to its P-rep: a supporting plane $\bm{p}_{t, sp}$ surrounded by three bounding planes $\{\bm{p}_{t, b}^i|\ i = 0,1,2\}$. We integrate the plane-based geometry into each of the tree steps of M\"{o}ller's algorithm as following:

\begin{figure}[t]
\centering
\includegraphics[width=3.5in]{sign}
\caption{We denote the signed distance from point $v_i$ to plane $\bm{p}_{t_2, sp}$ as $d_i$. All the four conditions of intersection between $t_1$ and $\bm{p}_{t_2, sp}$ are:  (a) $d_0\cdot d_2<0$, $d_1\cdot d_2<0$; (b) $d_0=0$, $d_1=0$, $d_2\neq 0$; (c) $d_0=0$, $d_1\cdot d_2<0$; (d) $d_0=0$, $d_1\cdot d_2>0$. End points of $Seg_1$ are intersections between $\bm{p}_{t_2, sp}$ and related edges of $t_1$ (red bold lines).}
\label{fig:isect}
\end{figure}

\begin{itemize}[leftmargin=0.45cm]
  \item[1)] The basic subroutine of the first step is given two faces $t_1$ and $t_2$, computing the orientation of of a face with respect to vertices from the other face. Reviewing the discussion of \S \ref{sec:convert}, under the constraint of $M \ge P+1$, we can compute the predicate within the precision of $M$ bits. This allow us to perform this early rejection efficiently.
      \vspace{0.5em}
  \item[2)] If $t_1$ and $t_2$ are not coplanar, the end points of $Seg_1$ and $Seg_2$ can be implicitly represented by plane triples. Take $Seg_1$ as an example. $Seg_1$ is the intersection between $t_1$ and $\bm{p}_{t_2, sp}$ and the end points of $Seg_1$ are intersections between $\bm{p}_{t_2, sp}$ and edges from $t_1$ (denoted as $\bm{e}^i_{t_1}$). Because $\bm{e}^i_{t_1}$ can be represented as $\bm{p}_{t_1, s}\cap \bm{p}^i_{t_1, b}$, the end points of $Seg_1$ can be represented in the form of $\bm{p}_{t_2, sp} \cap \bm{p}_{t_1, sp} \cap \bm{p}^i_{t_1, b}$. In Fig. \ref{fig:isect}, we list all possible intersecting situations between $t_1$ and $\bm{p}_{t_2, sp}$. The two intersection edges $\bm{e}^i_{t_1}$ for each situation are also presented. The coplanar situation will be discussed later in \S \ref{sec:degenerate}.
      \vspace{0.5em}
 \item[3)] The intersection between $t_1$ and $t_2$ is the overlap area of $Seg_1$ and $Seg_2$. It can be easily computed by sorting the end points of $Seg_1$ and $Seg_2$ along the line $\bm{l}\colon \bm{p}_{t_1, sp} \cap \bm{p}_{t_2, sp}$. Because end points are all represented using plane triples, we can use the linear order of points discussed in \S \ref{sec:substrates} to sort.
\end{itemize}

Each time an intersection is computed, two intersections are generated (one for each triangle) and the newly generated vertices are added into meshes. To avoid repetitive vertices in the final result, we perform vertex repetition elimination in this stage, merging coincident vertices even if they come from different input meshes. Because we use P-reps, the comparison of vertices is exact.

\subsection{Plane-Based Intersection Representation}
\label{sec:ir}

In our method, intersection between two triangles is denoted as $\bm{\mathcal{I}}\colon\{t, \bm{p}_{ext}, (\bm{p}^0, \bm{p}^1),\bm{\Psi}\}$, which is called plane-based intersection representation (PBI-rep). The first component $t$ indicates $\bm{\mathcal{I}}$ is on triangle $t$, and second one $\bm{p}_{ext}$ indicates $\bm{\mathcal{I}}$ lies on $\bm{p}_{ext}$, which also means it lies on line $\bm{p}_{t, sp} \cap \bm{p}_{ext}$. The two end points of $\bm{\mathcal{I}}$ are $\bm{p}_{t, sp} \cap \bm{p}_{ext}\cap\bm{p}^0$ and $\bm{p}_{t, sp} \cap \bm{p}_{ext}\cap\bm{p}^1$. The last component $\bm{\Psi}$ is a set that represents the neighborhood of the intersection. If $\bm{\mathcal{I}}$ is on the surface of mesh $P_k$, an element $(P_{k}, \Psi_{k})$ is added to $\bm{\Psi}$, where $\Psi_{k}$ is the edge or face of $P_k$ where $\bm{\mathcal{I}}$ lies.

\begin{figure}[t]
\centering
\includegraphics[width=2in]{pbi}
\caption{Triangle $t_1$ and $t_2$ are from mesh $P_i, P_j$ respectively. They intersect on line segment, generating intersection $\bm{\mathcal{I}}_{12}$ on $t_1$ and $\bm{\mathcal{I}}_{21}$ on $t_2$. The PBI-rep of $\bm{\mathcal{I}}_{12}$ is $\{t_1, \bm{p}_{t_2, sp}, (\bm{p}_{t_2, b}^0, \bm{p}_{t_2, b}^1),\{(P_i, \Psi_i), (P_j, \Psi_j)\}\}$.}
\label{fig:pbi}
\end{figure}

For example, in Fig. \ref{fig:pbi}, two triangle faces $t_1$ and $t_2$ from mesh $P_i$ and $P_j$ respectively intersect on a line segment. Then it generates two intersections for each of the two triangles, denoted as ${\bm{\mathcal{I}}}_{12}$ and ${\bm{\mathcal{I}}}_{21}$. For ${\bm{\mathcal{I}}}_{12}$, the first two components are $t_1$ and $\bm{p}_{t_2, sp}$. The computation of the third component has already been discussed in \S\ref{sec:embed}. The neighborhood $\bm{\Psi} = \{(P_i, \Psi_i), (P_j, \Psi_j)\}$. In most time, $\Psi_i=t_1$ and $\Psi_j=t_2$. However, sometimes, intersections may lie on the primitive edges instead of faces. In that case, $\Psi_i$ and $\Psi_j$ are edges (Fig. \ref{fig:twin}). We postpone this discussion until \S\ref{sec:degenerate}. The computation of ${\bm{\mathcal{I}}}_{21}$ is similar.

\subsection{Degenerate Case Handling}
\label{sec:degenerate}

While two triangles, if intersecting, intersect on a line segment in most situations, they can also intersect on a point or a convex area (coplanar case). Even if the intersection is a line segment, the intersection can be on triangle edges. These degenerate situations prevent us from performing robust boolean operations. In this section, we offer simple but effective way to deal with all these degenerations and conceal the complexity of intersections behind the concise PBI-rep.

%The criterion of whether an intersection help tessellation is by its necessity: if the intersection is not presented, whether some faces from the linked halfedge will cross the boundary of primitives. The word 'cross' do not only include the situation that a face is partly outside and partly inside of a primitive, but also can mean that a face is partly inside (outside) and part on the boundary of the primitive.
\begin{figure}[t]
\centering
\includegraphics[width=3.5in]{isolated}
\caption{Two situations of point intersections. (a) $\bm{v}_{iso}$ is isolated, and there is no adjacent intersection line segment. We need to add the $\bm{v}_{iso}$ to triangle $t_1$. (b) $\bm{v}_{non-iso}$ is not isolated. We do not need to add $\bm{v}_{non-iso}$ into $t_1$ because it can be added by other triangle pairs which intersect on line segments. However, for simplicity, we add the intersection point into $t_1$ in both situations.}
\label{fig:isolated}
\end{figure}

\subsubsection{Point intersection}
\label{sec:ipoint}
If two triangles intersect on a single point (Fig. \ref{fig:isolated}), the intersection cannot be represented using our four-component description. Therefore, we simply add the intersection point into the related triangles to guarantee correct tessellation, and no intersection is introduced.

\subsubsection{Edge intersection}

% intersect on face edges does not create new line segments, it is part of the edges from other meshes, and will to be duplicated.

% whose edge? need to be further explained and better defined. an edge intersection divide edge in a triangle and will be duplicated in the other

When intersection line segment lies on an edge (Fig. \ref{fig:twin}). we call it \emph{edge intersection}. The only difference between edge intersection and the general case is the neighborhood representation $\Psi$. The space near edge intersection is decided by both face in the two side of the intersected edge, instead of just one triangle face (Fig. X). Therefore, supposing intersection is on the edge $\bm{e}_{k, i}$ of mesh $P_k$, then the neighborhood $\Psi_k = \bm{e}_{k, i}$. Note that an intersection can be on edges from different meshes. Then there are more than one component of $\bm{\Psi}$ are edges (Fig. \ref{fig:twin}d).

Another thing needs to be noticed is that there will be repetitive detection of edge intersections. In Fig. \ref{fig:twin}a, the same intersection on $t_1$ will be detected twice with both $t_2$ and $t_2^{\star}$. We solve this duplication together with other interactions among intersections in \S\ref{sec:tessellation}. We call such $t_2$ and $t_2^{\star}$ as \emph{companion triangles} in an edge intersection, a concept which will be referred again in the discussion of coplanar cases.

\begin{figure}[t]
\centering
\includegraphics[width=3.5in]{twin}
\caption{Six situations of edge intersections. The first three subgraphs are edge-face intersections. The last three ones are edge-edge intersections. (a) Triangles are located in two sides, cross condition. (b) Triangles are located in the same sides, invalid condition. (c) One of the triangles are coplanar, coplanar condition. (d) Double twin intersections, cross condition. (e) Double twin intersections, coplanar condition. (f) Double twin intersections, coplanar condition.}
\label{fig:twin}
\end{figure}

\subsubsection{Copalnar}

\begin{figure}[t]
\centering
\includegraphics[width=3.5in]{coplanar}
\caption{a) Coplanar cases and the b) the possible configurations of the companion faces. In any configurations, their is no need to record the intersection of coplanar cases.}
\label{fig:coplanar}
\end{figure}

Consider two triangle faces $t_1$ and $t_2$ intersect within a common plane. Both $t_1$ and $t_2$ will divide each other into two areas---a convex overlapping area and exclusive parts (Fig. \ref{fig:coplanar}). Apparantly, if we tessellate both $t_1$ and $t_2$ according to the boundary of overlapping area, we can guarantee that the tessellated meshes will not cross the boundary of any other meshes. In fact, many methods \cite{feito2013fast,zhou2016mesh} claim that tessellating faces according to this 2D-intersection result is necessary. However, in our method, we do not test whether $t_1$ and $t_2$ really intersect in 2D. We treat them as they do not intersect at all. In this way, we simplify our method, making it more robust and fast, while doing no harm to the topology correctness.

The key observation is that intersection in 2D does not create new valid edges. The word `valid` means the edge is necessary or possibly necessary for the final result mesh. We find that each intersection line segment in 2D is on the edges from input meshes. That means we can view 2D intersection as a special case of edge intersection. The only difference is that there can be up to three edge intersections in one 2D intersection (Fig. \ref{fig:coplanar}). As we have discussed, edge intersection will be detected twice. That means we can rely on the companion triangle faces to detect intersections. However, if the companion triangles are also coplanar, neither of the two triangles will record the intersection. In fact, we do not need to worry about this, because in this case, the intersection is not valid. The two sides of the intersection are both retained or both dropped and we do not need it for tessellation.

\section{Deferred Tessellation}

\label{sec:tessellation}
After intersection computation, tessellation is performed on each intersecting triangle face and generate our intersection-free meshes. We call this stage as deferred tessellation because the tessellation happens until all valid intersections are computed.

In previous work, Ogayar-Anguita et. al. \cite{ogayar2015deferred} use Constrained Delaunay Triangulation (CDT) to perform deferred tessellation. They treat triangle faces as convex triangulation zones and intersections as constraints. Our first try is to implement a robust plane-based CDT. However, CDT algorithms \cite{chew1989constrained,preparata2012computational} usually require coordinates projection or intersection detection between arbitrary connections of vertices within triangulation zone. Because vertices may be implicitly represented in our method, these geometry predicates are hard to guarantee computation efficiency and precision simultaneously. Moreover, when there are more than two primitives, simply applying CDT omits the complexity of intersections between meshes. The intersections may intersect each other, introducing new vertices and splitting original intersection line segments (Fig. \ref{fig:iisect}). This breaks the assumption of most CDT algorithms and may lead to incorrect output topology.

\begin{figure}[t]
\centering
\includegraphics[width=2in]{iisect}
\caption{suppose? actual?}
\label{fig:iisect}
\end{figure}

To solve these problems, we first perform intersection refinement to eliminate any crossing or overlapping between intersections. After that, we perform conservative tessellation based on \emph{tess-graph}, ensuring unconditional topology correctness. An tess-graph is a graph-like description of intersections on a certain face. The word `conservative` means we tessellate face into polygons, not triangles. This is because if triangles are required, we need construct new bounding planes for these new triangles, which violating our principle of no geometry construction (Fig. x).

\subsection{Intersection Refinement}

The intersection vertices are not only introduced by intersection between an edge and a triangle, but also introduced by intersections of three triangles. It is not enough to find all of them by only triangle-triangle intersection tests. Also, because of degenerate cases, there will be other forms of intersections involving three or even more triangles. To make tessellation easier and more robust, we have to refine these intersections and all unusual conditions have to be removed before it.

Intersection refinement is performed in a local scope. For each intersecting face $t_i$, we collect all intersections on $t$ as a set $\bm{\Gamma}(t)=\{\bm{\mathcal{I}}_{t}^i, i=0,1,2,...\}\cup\{\bm{\mathcal{I}}_{t}^{e_i},i=0,1,2\}$ and refine them together. The set $\{\bm{\mathcal{I}}_{t}^{e_i},i=0,1,2\}$ is the PBI-reps of the three edges of $t$. We treat edges as intersections for simplicity in this stage. The neighborhood component $\bm{\Psi}(\bm{\mathcal{I}}_{t}^{e_i})$ is set as $\bm{0}$. Other components of $\bm{\mathcal{I}}_{t}^{e_i}$ are straightforward to determine. Refinement on $\bm{\Gamma}(t)$ is done using only plane-based geometry predicates by the following three steps:

\vspace{0.5em}
\noindent \textbf{Collinear splitting}~~~~
Each intersection $\bm{\mathcal{I}}_{t}^x$ are testing whether they are collinearly overlapping with other intersections. If yes, $\bm{\mathcal{I}}_{t}^x$ is split by the end points of overlapping intersections (Fig. x). Because $\bm{\mathcal{I}}_{t}^x$ may collinearly overlap more than one other intersection, the splitting may be deferred until all splitting points are found. The subroutine linear order of points (\S\ref{sec:substrates}) is used to sort these splitting points along $\bm{\mathcal{I}}_{t}^x$. The PBI-reps of split segments inherit their father's except the second component the end points.

\vspace{0.5em}
\noindent \textbf{Coincidence elimination}~~~~
We merge coincidence intersections which have the same end points. Intersections are undirected so intersections with inverse end points are also coincident. The PBI-rep of the new intersection is inherited from either of the old one's except the neighborhood component $\bm{\Psi}$. We follow the merging rules below:
\begin{equation}
  \Psi_k(\bm{\mathcal{I}}_{t}^{12})=\left\{
   \begin{array}{l}
   \Psi_k(\bm{\mathcal{I}}_{t}^1)\mbox{ or }\Psi_k(\bm{\mathcal{I}}_{t}^2),\ \mbox{ if } \Psi_k(\bm{\mathcal{I}}_{t}^1)=\Psi_k(\bm{\mathcal{I}}_{t}^2) \\
   \Psi_k(\bm{\mathcal{I}}_{t}^1),\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \mbox{ if } \Psi_k(\bm{\mathcal{I}}_{t}^2)=0  \\
   \Psi_k(\bm{\mathcal{I}}_{t}^2),\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \mbox{ if } \Psi_k(\bm{\mathcal{I}}_{t}^1)=0  \\
   \end{array}
  \right.
\end{equation}
, where $\bm{\mathcal{I}}_{t}^1$ and $\bm{\mathcal{I}}_{t}^2$ are old intersections and $\bm{\mathcal{I}}_{t}^{12}$ is the new one.

\vspace{0.5em}
\noindent \textbf{Crossing intersection resolving}~~~~
After the process of two previous steps, there are no overlapping intersections in $\bm{\Gamma}(t)$. We check whether any pair of intersections intersect each other. The definition of \emph{intersect} does not include the situation that two intersections share end points. In Fig. x, we illustrate the only two situations. In both situations, one intersection is split by other, and the same splitting routine is used as in collinear splitting.

%\vspace{0.5em}
%There is one more thing that we give intersections generated (by merging or splitting) from the PBI-reps of edges  the positive directions. That means all intersections lies on triangle edges have a positive direction. Direction is defined by the direction of edge the intersection lies on.

\subsection{Tessellation by Tess-Graph}
\label{sec:tess}
So far, intersections are only independent data that store the coordinates and neighbouring information. To perform tessellation and extract subfaces, we need to organize them to reveal the topology. We use tess-graph for this purpose. A tess-graph is the graph description of the tessellated face topology. For each face to be tessellated, we construct a tess-graph according to the refined intersections. Nodes of tess-graph represent end points of intersections and undirectional connections between nodes represent intersections. The construction of tess-graph is straightforward and the reader will not have problem filling the details.

After we have tess-graph, we tessellate face and extract subfaces according to it. This is done by extract valid loops on tess-graph. A valid loop that corresponds to a subface must satisfy two criterions: 1) the direction of loops should be coherent with the face normal, and 2) consecutive connections on the loop should be adjacent by circular order ({\color{red}{Fig.?}}). Each valid loop corresponds to a intersection-free face. After we find out all valid loops on tess-graph, we finish tessellating the corresponding face.

When all faces are tessellated, meshes are intersection-free. We further attach the PBI-reps of intersections to the edges generated from tess-graph connections which corresponds to an intersection. Finally, we get the intersection-free hybrid meshes, while `hybrid` means some parts of meshes are both vertex-based and plane-based represented.

%In our implementation, we find that it is not easy to check whether the loop direction is coherent with face normal by plane-based geometry.

%First, We classify connections of tess-graph into two types: connections on face boundaries as \emph{boundary connections} and connections inside face as \emph{inner connections} by the whether they have positive directions. In the view of halfedge, inner connections represent two opposite halfedges but boundary connections represent only one. We add an attribute for each connection to identify their directions, and thus we get a directed version of Tess-graph. A sub-face is a halfedge loop on the directed Tess-graph. Extracting faces are process of searching for valid loops which meet certain geometry constraints. After all faces are extracted, all halfedges of directed Tess-graph should be used

%There two geometry constraints for valid loops. Second, consecutive connections on the loop should be adjacent by circular order {\color{red}{Fig.?}}. To meet these constraints, we sort the halfedges counter-clockwise for each nodes with more than two halfedges starting from them. This is implemented by using the divide-and-conquer framework of quick sort. During each recursion, we pick a random halfedge and split the rest into two sets according to the sign of two-line-on-plane comparison. To make the results of two-line-on-plane comparisons coherent, we assume that for each halfedge $h_i$, the normal directions of its plane-rep, denoted as $\vec{n}(h_i)$, meets the following criterion:
%\begin{equation}
%(\vec{n}(h_i) \times \vec{n}_0) \cdot \vec{l}(h_i) > 0,
%\end{equation}
%where $\vec{n}_0$ is the normal of the face, and $\vec{l}(h_i)$ is the direction of $h_i$. The plane-rep of halfedge can be inherited from the corresponding connections from the undirected Tess-graph. Since we {\color{red}{guarantee that each connection has a plane-rep //need insert//}}, sorting can be exactly and fast performed. After connections are sorted in each nodes, we can know which connections are adjacent and what is the right direction during the loop search.

\section{Exact Face Classification}

\label{sec:classification}
We traverse each face in the hybrid meshes, and classify which one belongs to the final mesh. Classification is done by evaluating the indicator vector of each face. The basic idea of our classification method is to utilize the space coherence of face indicator vectors to reuse the classification results. Also, because we store the PBI-reps which contains neighborhood information in the edges which intersect other meshes, the search space of point-in-polyhedron test is limited to only a few faces, which makes the indicator computation very fast. Also, by using P-reps, our classification method can deal with any degenerate cases.

\subsection{Framework}

The space coherence of indicator vectors means neighboring faces could share the same indicator vector, or most components of indicator vectors. Therefore, we start from a seed face $s_0$ with its indicator vector $\bm{\Lambda}(s_0)$ known, and propagate to other faces in a flood-filling manner. The trace of indicator propagation is:
\begin{equation}
\label{eq:trace}
s_0\to s_1\to s_2\to s_3\to ...
\end{equation}

Our hybrid meshes facilitate this process by accelerating the indicator propagation. Given adjacent faces $s_1$ and $s_2$, it is straightforward to know whether $\bm{\Lambda}(s_1)$ and $\bm{\Lambda}(s_2)$ are different. The two indicator vectors are the same unless the shared edge $e_{12}$ lies on the boundary of some meshes, and it means there is an intersection $\bm{\mathcal{I}}(e_{12})$ stored on $e_{12}$. For a certain indicator $\lambda_k$, it always stands:
\begin{equation}
\label{eq:eq1}
{\lambda_k}(s_1) = {\lambda_k}(s_2),\ \ \  \mbox{if }\Psi_k(\bm{\mathcal{I}}(e_{12}))= 0.
\end{equation}
Even if $\Psi_k(\bm{\mathcal{I}}(e_{12})) \neq 0$, both $\lambda_k(t_1)$ and $\lambda_k(t_2)$ can be computed efficiently and exactly according to $\bm{\mathcal{I}}(e_{12})$ (\S\ref{sec:propagation}). We outline our classification method in Algorithm \ref{code:floodfill}.

\begin{algorithm}
\caption{Fast Face Classification}
\label{code:floodfill}
\textbf{Input: } IFH meshes $\bm{P'}$, a boolean function $f$

\textbf{Output: } Classification $f(\bm{\Lambda}(s_i))$, for all faces $s_i$ from $\bm{P'}$

\begin{algorithmic}[1]
\State Select a proper seed face $s_0$
\State Compute the first indicator vector $\boldsymbol{\Lambda}(s_0)$
\State \Call{propagate} { $s_0$ , $\boldsymbol{\Lambda}(s_0)$}
\State
\Function{propagate}{ $s$ , $\boldsymbol{\Lambda}(s)$}
    \State Compute $f(\boldsymbol{\Lambda}(s))$
    \For {each neighboring face $s_i$}
        \If {$s_i$ has been classified}
            \State continue
        \EndIf
        \If {there are PBI-reps $\{{\bm{\mathcal{I}}}_k\}$ on $e(s_i, s)$}
            \State compute $\boldsymbol{\Lambda}(s_i)$ by $\boldsymbol{\Lambda}(s)$ and $\{{\bm{\mathcal{I}}}_k\}$
            \State \Call{propagate} { $s_i$ , $\boldsymbol{\Lambda}(s_i)$}
        \Else
            \State \Call{propagate} { $s_i$ , $\boldsymbol{\Lambda}(s)$}
        \EndIf
    \EndFor
\EndFunction
\end{algorithmic}
\end{algorithm}

\iffalse
\subsection{Precise Seed Generation}

Our classification method starts from computing the seed indicator vector using point-in-polyhedron test \cite{ogayar2005point}. Conventionally, the barycenter of the face is used for the test to compute the indicator vector of the whole face. This is because the whole face is classified as a whole, and every inner point will have the same indicator vector. However, the coordinates of the inner points cannot be exactly represented by float point numbers, which may lead to wrong predicates, especially when it is an on-boundary case. Therefore, we use the original vertices, whose exact coordinates are known, for point-in-polyhedron test. For this reason, only tessellated faces with at least one original vertex can be the seed face. The point-in-polyhedron test is guaranteed to be exact. On the other hand, efficiency is not so important as there will not be many such tests. If ray-trace algorithms (e.g., \cite{frisken2002simple}) are used, the octree constructed during space division can be used for acceleration \cite{havran1999summary}.

However, using point on the boundary instead of an inner one has another problem. Given a face $s$ (may not be triangle) and one of its vertex $v_b^i(s)$, the vertex indicator $\boldsymbol{\Lambda}(v_{i}(s), P_k)$ may cause ambiguity to deduce the face indicator $\boldsymbol{\Lambda}(s, P_k)$. In {\color{red}{Fig. ?}}, $\boldsymbol{\Lambda}(v_b^1(s), P_k) = on$, then $\boldsymbol{\Lambda}(s, P_k)$ can be any of the four conditions. We solve this problem by trying other vertices in $s$ for ambiguous indicators. If all vertices indicators are $on$ and the ambiguity cannot be resolved, the current face is not proper for being the seed and we select a new one instead.

\fi

\subsection{Indicator Propagation}
\label{sec:propagation}
Now, our meshes consist of vertices, edges and faces. Classification is required to perform on face level. However, polyhedron in-out test is performed based on points. Conventionally, face inner point such as barycenter is used to compute the indicator of the whole face. This is because in intersection-free meshes, face is classified as a whole, and every inner point will have the same indicator as the face's. However, the coordinates of inner points cannot be exactly represented by float point numbers, which may lead to wrong predicates, especially when the face lies on the surface of other mesh (Fig. x). Therefore, we use the vertices, whose exact representations are known, for indicator computation.

%For example, given a face $s$ and one of its vertex $v^i(s)$, if $\lambda_k(v^i(s))=on$, $\lambda_k(s)$ can be any of $on$, $in$ or $out$ (see {\color{red}{Fig. ?}})
However, there are two gaps between face indicator and vertex indicator. First, the indicator of vertex does not always equal to indicator of face. Second, face indicator has two conditions in $on$ case, $same$ and $oppo$, because the face has its orientation.

To solve these problems, we discuss our indicator trace (\ref{eq:trace}) in finer granularity. In the beginning, we start from a seed vertex $v_0$ on seed face $s_0$. We assume indicator vector $\bm{\Lambda}(v_0)$ is known. Then we use $\bm{\Lambda}(v_0)$ to compute $\bm{\Lambda}(s_0)$. In addition, we add the edge layer into the trace. Then it changes to:
\begin{equation}
v_0\to e_0\to s_0\to e_1\to s_1\to e_2\to s_2\to ...
\end{equation}
, where $v, e$ means vertex and edge respectively. We find that there are three kinds of basic operations in the trace: $v_e\to e$, $e_s\to s$ and $s\to e_s$, where $v_e$ means the vertex is the end point of $e$, and $e_s$ means the edge is the edge of $s$.

Given the partial order $on \succ in$ and $on \succ out$, our key observation is that the following relation always stands for indicators:
\begin{equation}
\label{eq:porder}
\lambda_k(v_{e_s}) \succeq \lambda_k(e_s) \succeq \lambda_k(s)
\end{equation}
, where $\lambda_k(x)$ means the indicator of $x$ against a certain primitive $P_k$. This relation can be inferred from continuous space assumption and the definition of intersection-free mesh (def. \ref{eq:ifmdef}). It tells us when we perform $v_e\to e$ and $e_s\to s$, we decide whether the $on$ indicator change to $in$ or $out$ or remain $on$. When we perform $s\to e_s$, we decide whether $in$ and $out$ indicators change to $on$. In addition, considering the orientation of face, we also need to figure out whether the $on$ indicators change to $same$ or $oppo$ during $e_s\to s$ operation.

\vspace{0.5em}
\noindent\textbf{$\bm{s\to e_s}$}~~~~ When $\lambda_k(e_s)=on$, there will be PBI-rep $\bm{\mathcal{I}}(e_s)$ attached to $e_s$ and the related neighborhood $\Psi_k(\bm{\mathcal{I}}(e_s))$ must be nonzero, since it is on the surface of mesh $P_k$:
\begin{equation}
\label{eq:eq1}
\lambda_k(e_s)=on,\ \ \  \mbox{iff. }\Psi_k(\bm{\mathcal{I}}(e_s))\neq 0.
\end{equation}
So all we need is to iterate over all nonzero components of $\bm{\Psi}(\bm{\mathcal{I}}(e_s))$ and change corresponding indicators to $on$.

\vspace{0.5em}
\noindent\textbf{$\bm{e_s\to s}$}~~~~From eq. \ref{eq:porder}, we know if $\lambda_k(e_s) \neq on$, then $\lambda_k(e_s)=\lambda_k(s)$. When $\lambda_k(e_s) = on$, we figure out $\lambda_k(s)$ using $\Psi_k(\bm{\mathcal{I}}(e_s))$. From eq. \ref{eq:eq1}, we know $\Psi_k(\bm{\mathcal{I}}(e_s)) \neq 0$ and can be either a triangle face or an edge. If $\Psi_k(\bm{\mathcal{I}}(e_s))$ is a face, we can build a trivial BSP \cite{thibault1987set} using the face on the space near $e_s$. If $\Psi_k(\bm{\mathcal{I}}(e_s))$ is an edge, we can still build BSP using the opposite two faces along $e_s$. In both cases, the BSP can be used to compute $\lambda_k(s)$ if we can sample a point from $s$ that near $e_s$. Unfortunately, we cannot guarantee to find such a point with precise coordinates. Instead, we choose a point $v_x$ from the half plane $\bm{p}_{s, sp}^+$, which is on the supporting plane of $s$ and on the positive side of bounding plane $\bm{p}_{s, b}^{e_s}$ (Fig. x). For polygons, there is at least one such $v_x$ among the vertices of $s$. We find that the BSP is so simple that indicators $\lambda_k$ of all points on $\bm{p}_{s, sp}^+$ are the same. Since $\bm{p}_{s, sp}^+ \cap s \neq \varnothing$, we can compute $\lambda_k(s)$ by figuring out $\lambda_k(v_x)$. Also, if we assign each BSP splitting plane a normal by the normal of triangle it contains, we can use it to decide the orientation ($same$ or $oppo$) when $v_x$ is on a splitting plane, which means $\lambda_k(s)=on$.

\vspace{0.5em}
\noindent\textbf{$\bm{v_e\to e}$}~~~~ While this is yet another propagation from low dimension to high dimension, there is no neighborhood information stored on vertex. If $\lambda_k(v_e)=on$, we first check whether $\lambda_k(e) = on$ by (\ref{eq:eq1}). If not, we need to find where $v_e$ lies on the surface of $P_k$ and determine the neighborhood of $v_e$. In most cases, it is straightforward since coincident vertices among different intersection-free meshes are shared. Exceptions are discussion is in \S\ref{sec:esubroutine}. The neighborhood may be a vertex, an edge or a face. After computing the neighborhood, the same BSP-based classification method is used as in ${e_s\to s}$ to determine the high dimensional $\lambda_k(e)$. We use the other vertex of $e$ to be the sample point $v_x$ for the BSP classification.

\iffalse
\subsection{Implementation}

\label{sec:individual}

Feito et. al. \cite{feito2013fast} had noticed that intersection neighborhood can be used for fast classification of faces around that intersection. According to their method, faces adjacent to intersection have neighbors from other primitive(s) that define the position and orientation for each other, by which the indicators are determined. However, Feito et. al. did not give detail description of how to implement an exact and robust classification. Their implementation of using vertex coordinates might misclassify faces by numeric errors and fail in degenerate cases. Because geometry connectivity is used, local misclassification can be spread to the neighborhoods, leading to errors in wide ranges.

We develop an classification method that is unconditionally robust and exact based on the idea of using intersection neighborhood information. The neighborhood information is stored in the context component of I-reps during previous stages . In the following, we discuss the conditions that the face to be classified is a triangle. If the face is not a triangle, we select a triangle slice of the face instead, as the whole face is guaranteed to be classified together. Note that the chosen slice should contain the edge where the I-reps lie, by which we compute the new indicators.

Given a face $t_i$ to be classified and an I-rep $\boldsymbol\Gamma$ of intersection on one of $t_i$'s edge, we denote the three corners it as $v_{b*}^{0}(t_i), v_{b*}^{1}(t_i), v_b^2(t_i),$ where the subscript $*$ means the vertices are on the edge where $\boldsymbol\Gamma$ lies. We also denote the context component of $\boldsymbol\Gamma$ as $C_{P_h \backslash P_j}$, meaning $t_i$ is from primitive $P_j$ and intersects primitive $P_h$. From previous discussion (section \ref{sec:ir}){\color{red}{may be more than one}}, we know there are two types of intersection context---face neighborhood and edge neighborhood. In both conditions, the one or two faces involving in the context split the space into two in and out space by its (their) orientation(s). We will see that under such division (denoted as $\boldsymbol{D}(C_{P_h \backslash P_j})$), we can compute $\lambda(t_i, P_j)$ exactly.

Since the faces involving in $C_{P_h \backslash P_j}$ is part of primitive $P_j$, the space division $\boldsymbol{D}(C_{P_h \backslash P_j})$ is homogeneous with the space division by $P_j$ in neighborhood of $\bm{\mathcal{I}}$. The basic idea is to choose one point $v_x(t_i)$ on $t_i$ close enough to $\bm{\mathcal{I}}$ to compute $\lambda(v_x(t_i), P_j)$, and then compute $\lambda(t_i, P_j)$ accordingly. However, the problem is that we cannot find such a $v_x(t_i)$ that can be represented exactly, which means robust classification cannot be performed. Fortunately, we observe that we can choose $v_b^2(t_i)$, which is not so close to $\bm{\mathcal{I}}$ instead. To illustrate, assume there is another point $v_{x'}(t_i)$ inside of $t_i$ that is very close to $v_b^2(t_i)$ ({\color{red}{Fig. x}}). Because all inner points of $t_i$ have the same indicator, $\lambda(v_x(t_i), P_j) = \lambda(v_{x'}(t_i), P_j)$. On the other hand, we find that $\lambda(v_{x'}(t_i), P_j)$ and $\lambda(v_b^2(t_i), P_j)$ are always equal under division $\boldsymbol{D}(C_{P_h \backslash P_j})$ ({\color{red}{see Appendix \ref{app:profind}}}). Therefore, $\lambda(v_b^2(t_i), P_j)$ is equal to $\lambda(v_x(t_i), P_j)$ under $\boldsymbol{D}(C_{P_h \backslash P_j})$, and are coherent with $\lambda(t_i, P_j)$ under the division of whole primitive $P_j$. One additional thing is that when $\lambda(v_2(t_i), P_j)=on$, then $\lambda(t_i, P_j)$ can be either $same$ or $oppo$. The normal orientation test have to be performed to eliminate such ambiguity.

\fi

\section{Discussion}

\vspace{0.5em}

\subsection{Essential Subroutines}
\label{sec:esubroutine}

\vspace{0.5em}
\noindent\textbf{Searching valid loops}~~~~ In \S\ref{sec:tess}, we claim a valid loops on tess-graph has to be correct with its orientation. However, since tess-graph is undirectional, it is not easy to check whether a loop orientation agrees with the face's. We observe that the graph connections lie on face edges have their unambiguous directions (Fig. x). Starting from these edges, we can guarantee that the loops have the right orientations. After we find a valid loop, more graph connections will have unambiguous directions because one connection can participate two valid loops at most. In such way, the correct orientation will propagate in a flood-filling way.

\vspace{0.5em}\noindent There is another problem that the tess-graph may not be a connected graph. We avoid this problem by inserting fake-intersections into the tess-graph. An fake intersection links the vertex of face (on the outer connected component) with the vertex on the inner connected component. The intersection refinement has to be performed again if the fake intersections cross any other intersections. To guarantee the fake intersection has a valid PBI-rep, we require the vertex from inner connected component generated by edge-face intersection, which means it is introduced during triangle-triangle intersection test. In this way, we can find an exact plane with the fake intersection on it (Fig. x).

\vspace{0.5em}
\noindent\textbf{Seed indicator generation}~~~~In \S\ref{sec:propagation}, we claim that our flood-filling starts from a seed vertex $v_0$. While the indicator vector of the seed $\bm{\Lambda}(v_0)$ can be generated by point-in-polyhedron test \cite{ogayar2005point} with the octree as acceleration structure \cite{frisken2002simple}, we have simpler strategies by using vertex with known indicators as the seed. We choose the `peak vertex` with the max x coordinate as the seed. In this way, $\lambda_k(v_0)$ is either $out$ ($in$, if complement is applied on mesh) or $on$. The $on$ indicators are easy to determine in most time. Though sometimes exception occur because of multi-coplanar situation, this will not happen if we carefully choose the right peak vertex whose neighbor faces are not coplanar (Fig. x).

\vspace{0.5em}\noindent The indicator vectors can propagate not only among single meshes, but can also between different meshes by the vertices and edges they share. Therefore in most time, we only need one single seed vertex for classification. However, if there are lonely meshes that do not intersect with other meshes, we need to new seeds. The indicators of the second and later seeds can only be computed by point-in-polyhedron tests.

%Our method of generating the first indicator vector will fail on extreme conditions when all faces in Linked halfedge are coincident with more than one primitive face (e.g., the union of the same mesh). In order to be robust, we use the high-precision face barycenter with the point-in-polyhedron test to generate the seed indicator vector instead after several failures of choosing a proper seed face.

\iffalse

\subsection{Robustness}

\vspace{0.5em}
\noindent\textbf{Indicator up-propagation}~~~~
In \S\ref{sec:propagation}, we point out that up-propagation (from low dimension to high dimension, $v_e\to e$ and $e_s\to s$, etc.) can be done by using sample points. Here we give use a heuristic method to prove our claim in $e_s\to s$. Extension to $v_e\to e$ should not take much effort.

\vspace{0.5em}\noindent We prove $\lambda_k(v_x)=\lambda_k(v_{\delta})$ in the BSP of $e_s$'s neighborhood, where $ v_x$ is the sample point from $\bm{p}_{s, sp}^+$ and $v_{\delta}$ is from the neighborhood $U(e_s, \delta)$. If $\lambda_k(v_x) \neq on$, then

Given $v_{x'}(t_i)$ can be arbitrarily close to $v_b^2(t_i)$, if $\lambda(v_b^2(t_i), P_j) = in | out$, the equality holds because of the continuity of space. If $\lambda(v_b^2(t_i), P_j) = on$, we divide the proof of equality into two parts according to the type of intersection context. If $C_{P_h \backslash P_j}$ is face neighboring type, $\boldsymbol{D}(C_{P_h \backslash P_j})$ is a single face, say, $t_0^c$. We have already know that the other two corners $v_{b*}^{0}(t_i)$ and $v_{b*}^{1}(t_i)$ are on $s_0$. Since $\lambda(v_b^2(t_i), P_j) = on$, we know $v_b^2(t_i)$ is on $t_0^c$, too. There comes the conclusion that the whole triangle $t_i$ is coplanar with $t_0^c$. Therefore, $v_{x'}(t_i)$ is on $t_0^c$ and $\lambda(v_{x'}(t_i), P_j) = on$, too. On the other hand, if $C_{P_h \backslash P_j}$ is edge neighboring type, $\boldsymbol{D}(C_{P_h \backslash P_j})$ are two neighboring faces (denoted as $t_1^c, t_2^c$). We know that both $v_{b*}^{0}(t_i)$ and $v_{b*}^{1}(t_i)$ are on the shared edge $e(v_{b*}^{0}(t_i), v_{b*}^{1}(t_i))$. It means both vertices are on both $t_1^c$ and $t_2^c$. Using the same deduction of the first condition, we can conclude $\lambda(v_{x'}(t_i), P_j) = on$ whenever $\lambda(v_b^2(t_i), P_j) = on$.

\vspace{0.5em}
\noindent\textbf{Proof of Robustness}

\fi

\subsection{Extensions}

\vspace{0.5em}
\noindent\textbf{Converting to float-point number}~~~~
The vertices of final mesh in our method are represented in both P-reps and plain coordinates. While the vertices from the input meshes have there exact coordinates, the vertices newly introduced by intersection between meshes require round-off when converting to coordinates representation. Though we guarantee the correct topology in the result mesh, such round-off can still cause topology deficiencies such as self-intersections. Here we adopt Zhou et al.'s method \cite{zhou2016mesh} to solve it iteratively.

\vspace{0.5em}
\noindent\textbf{Incremental evaluation}~~~~While we claim in many CSGs, a non-incremental evaluation is faster by avoiding repetitive computation, one can easily find counter-examples that non-incremental evaluation contains useless intersection computation. For example, if 100 spheres densely locate in a small region, each sphere will need to intersect with other 99 spheres. However, if evaluating incrementally, we do not need computing so many intersections. Fortunately, our method can also perform incremental evaluation. For exactness and robustness, we have to deal with non-triangle input meshes in incremental evaluations and intersection vertices should not be converted to float-point.

\vspace{0.5em}
\noindent\textbf{Extending to polygon meshes}~~~~
In our method, we claim the inputs are triangle meshes mainly to avoid plane fitting when converting to P-reps. However, if we assume the P-reps of non-triangle faces are exactly known, our method can be easily extend to polygon meshes.

\vspace{0.5em}\noindent The only different part is intersection computation. The first step of M\''{o}ller's triangle-triangle intersection method is to test each edge-plane pair respectively. This idea can be transferred to polygon intersections without any change. Then we can get two plane-face intersections $Seg1$ and $Seg2$, except that in polygon cases, $Seg1$ and $Seg2$ are two sets of line segments, not single segments. And the later processes are the same.

\vspace{0.5em}
\noindent\textbf{Classification acceleration}~~~~If the CSG tree is large with hundreds of primitive nodes, computing $\lambda_f(s_x)$ from $\boldsymbol{\Lambda}(s_x)$ for every face $s_x$ can be costly. Thanks to the indicator space coherence, we can save the computation time by caching the evaluation result $f(\boldsymbol{\Lambda}(s_x))$ during flood-filling.

The basic cache strategy is to cache final result, that is, the value of $\lambda_f(s_x)$. We know indicator vector will not change if there is no intersections during propagation. That means the final classification $\lambda_f(s_x)$ will not change, too. Thus, those faces sharing the same indicator vector can be classified as a whole.

Also, we can perform intermediate results cache. We noticed that the boolean expression can be simplified if some components of $\boldsymbol{\Lambda}(s_x)$ is fixed. For example, assume we have a boolean expression $f = P_1\cup (P_2\cap P_3-P_4)$. Given the values of two indicators $\lambda_1(s_x)=out$, $\lambda_2(s_x)=in$, the expression can be rewritten as $f(\lambda_1=out, \lambda_2=in)=out\cup (in\cap P_3-P_4)$. Using the combination rules we can simplify the expression as $f(\lambda_1=out, \lambda_2=in)=P_3-P_4$. This fact is important because for a large CSG, a certain primitive (denoted as $P_k$) often only intersects with a few other primitives $\Theta(P_k)= \{P_{n_1}, P_{n_2}, \cdots, P_{n_k}\}$. That means all the faces in $P_k$ has the same indicators $\lambda_{i}, P_{i} \notin \Theta(P_k)$. Therefore, we can first determine these fixed indicators and simplify the boolean function, and then use simplified one to compute the final indicator for each faces in $P_k$.

\iffalse
\vspace{0.5em}
\noindent\textbf{Lazy intersection}~~~~
Another way to avoid useless intersection computation is to use lazy intersection technique
\fi

\section{Experiments}

\section{Summary}

In this paper, we proposed a novel method to evaluate CSG with triangular mesh primitives. It is able to efficiently perform non-incremental evaluation of large CSG with massive faces. The key idea of our approach is to use the local coherence of face space labels to accelerate face membership classification. A two-level grouping framework is developed to group neighboring faces together and thus space labels can be shared within each group. This scheme saves much time for space label computation, which is often very time-consuming for conventional Boolean evaluation algorithms. Additionally, in order to be robust, plane-based geometry computation is introduced into the intersection computation step of our approach. Experiments have verified the proposed approach is more efficient than the state-of-the-art techniques while retaining robustness and stability. We will further investigate the robust tessellation in future.

\appendices

%\newpage
\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,citation}

%\newpage

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{rui}}]{Rui Wang}
is currently a postgraduate student at the Department of Computer Science and Technology, the Shanghai Jiao Tong University. His main research interests include real-time computer graphics and virtual reality applications.
\end{IEEEbiography}

% if you will not have a photo at all:
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{xudong}}]{Xudong Jiang}
received his Master degree in Computer Science from Shanghai Jiao Tong University in 2014. He is currently working in Autodesk China Research \& Development Center. His research interest includes computer-aided geometric design and solid modeling.
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{hongbo}}]{Hongbu Fu}
is an Associate Professor in the School of Creative Media, City University of Hong Kong. He received the PhD degree in computer science from the Hong Kong University of Science and Technology in 2007 and the BS degree in information sciences from Peking University, China, in 2002. His primary research interests fall in the fields of computer graphics and human computer interaction. He has served as an associate editor of The Visual Computer, Computers \& Graphics, and Computer Graphics Forum.
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{sheng}}]{Bin Sheng}
received his BS degree in computer science from Huazhong University of Science and Technology in 2004, MS degree in software engineering from University of Macau in 2007, and PhD Degree in computer science from The Chinese University of Hong Kong in 2011. He is currently an associate professor in the Department of Computer Science and Engineering at Shanghai Jiao Tong University. His research interests include virtual reality, computer graphics, and image-based techniques.
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{wu}}]{Enhua Wu}
received the BS degree from Tsinghua University in 1970, and the PhD degree from the University of Manchester (UK) in 1984. He is currently a research professor at the Institute of Software, Chinese Academy of Sciences, and Fellow of China Computer Federation. He has also been teaching at the University of Macau since 1997, where he is currently an Emeritus Professor. His research interests include realistic image synthesis, virtual reality, and scientific visualization. He has served as an associate editor of The Visual Computer, Computer Animation and Virtual Worlds.
\end{IEEEbiography}

\end{document}